*Prochlorococcus marinus* responses to light and oxygen

### 2024-06-03 Code/Manuscript_O2.Rmd
Doug text edits

*Prochlorococcus marinus*, the smallest picocyanobacterium, comprises
multiple clades with distinct niches, currently across tropical and sub-tropical
oligotrophic ocean regions, including Oxygen Minimum Zones. Ocean
warming may open permissive temperatures in new, poleward photic
regimes, along with expanded Oxygen Minimum Zones. We used ocean metaproteomic
data to guide testing of *Prochlorococcus marinus* growth across a matrix of peak irradiances,
photoperiods, spectral bands and dissolved oxygen. MED4 from Clade HLI
requires greater than 4 h photoperiod, grows at 25 µmol O~2~ L^-1^ and
above, and exploits high cumulative diel photon doses. MED4, however, relies upon an alternative oxidase to
balance electron transport, which may exclude it from growth under our
lowest, 2.5 µmol O~2~ L^-1^, condition. SS120 from Clade LLII/III is
restricted to low light under full 250 µmol O~2~ L^-1^, shows expanded
light exploitation under 25 µmol O~2~ L^-1^, but is excluded from growth
under 2.5 µmol O~2~ L^-1^. Intermediate oxygen suppresses the cost of
PSII photoinactivation, and possibly the enzymatic production of H~2~O~2~ in SS120,
which has limitations on genomic capacity for PSII and DNA repair. MIT9313 from
Clade LLIV is restricted to low blue irradiance under 250 µmol O~2~
L^-1^, but exploits much higher irradiance under red light, or under
lower O~2~ concentrations, conditions which slow photoinactivation of
PSII and production of reactive oxygen species. In warming oceans, range expansions and competition among clades will be governed not simply by light levels. Short photoperiods governed by latitude, temperate winter, and depth attenuation of light, will exclude Clade HLI (including MED4). In contrast, clade LLII/III (including SS120), and especially clade LLIV (including MIT9313), may actually exploit higher light niches under expanding OMZ conditions, where low O~2~ relieves the stresses of oxidation stress and PSII photoinhibition.

@@

Low-Light clades thrive in deeper ocean waters, extending beyond 200 m
in depth [@partenskyProchlorococcusMarinePhotosynthetic1999], where only
\~1% of the surface irradiance penetrates, primarily in the blue (450
nm) to green (520 nm) spectral range
[@holtropVibrationalModesWater2021]. Clade LLI includes cultured strain
NATL2A, which prefers moderate irradiances typical of between 30 and 100
m depth. Clades LLII and LLIII, including cultured strain SS120, are
grouped together as early branching phylogenetic lineage in
the *P. marinus* radiation, with a preference for low light. Clade
LLIV, including cultured strain MIT9313, falls near the base of the
*P. marinus* radiation, and has been characterized by preference for low
light, typical of depths from 120 m to 200 m
[@partenskyProchlorococcusMarinePhotosynthetic1999]. LLIV members are,
as yet, the only cultured *P. marinus* strains to have been found in Oxygen Minimum
Zones (OMZ). Some, as yet, uncultured *P. marinus* strains in clades LLV
and LLVI also thrive in OMZ of the subtropical Atlantic and Pacific
Oceans, where dissolved oxygen concentrations [O~2~] can be less than 20
µM [@ulloaCyanobacteriumProchlorococcusHas2021;
@lavinNovelLineagesProchlorococcus2010;
@goerickeNovelNicheProchlorococcus2000;
@garcia-robledoCrypticOxygenCycling2017;
@johnsonEnergeticsGrowthKinetics1999]. *Prochlorococcus marinus* LL
ecotypes may indeed dominate the phytoplankton within OMZ
[@lavinNovelLineagesProchlorococcus2010;
@ulloaCyanobacteriumProchlorococcusHas2021;
@partenskyProchlorococcusAdvantagesLimits2010], where they may be net
O~2~ consumers [@partenskyComparisonPhotosyntheticPerformances2018].

@@

## Detection of *Prochlorococcus marinus* proteins across O~2~ and light niches in the ocean {.unnumbered}

Proteins from 8 annotated strains of *P. marinus* were detected across
depths and oxygen concentrations in the OPP data set
analyzed. We focused our analysis here on core photosynthetic protein
complexes, for clades HLI (2 strains), LLI (2 strains), LLII/III (2 strains) and LLIV (2 strains) (Fig \@ref(fig:CladeProchloroPhotosynthDepthO2)) as a function of depth (a
proxy for light intensity) and measured [O~2~] at the sampling locations. Photosynthetic complexes
from HLI (including strain MED4) and LLI (including strain NATL2A) were detected throughout the water
column, predominately under high [O~2~]. Though present in the genome of MED4, the absence of proteins annotated as RUBISCO complex for clade HLI, compared to annotated
detections of RUBISCO across the other three clades, suggests
limitations in the annotation process for highly conserved protein
sequences. Notwithstanding limitations on assignments of proteins to clades, detections of the abundant carbon fixation complex RUBISCO derived from HLI, LLI and LLII/III were notably absent from samples taken at lower [O~2~], suggesting limited capacity for carbon fixation by HLI,  LLI and LLII/III under low [O~2~] habitats. Clade LLII/III (including strain SS120)
photosynthetic complexes were also detected throughout the depth/light and [O~2~] ranges, notably with the most frequent detections of proteins at both high and low [O~2~], compared to other strains. Complexes from LLIV (including strain MIT9313) were also detected across the depth/light and [O~2~] ranges, with more detections at deeper, darker depths and at low [O~2~].Our analyses utilized a proteomics dataset; however, alternate approaches, such as metagenomics or metatranscriptomics, could have been employed to analyze ecotype abundances using the TARA Oceans [@pesantOpenScienceResources2015] or other datasets.

@@

# This is Fig2.tiff
CladeProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections (circles) are plotted vs. O~2~ (µM) (x-axis) and depth (m) (y-axis) at sample origin with a 15% offset to seperate protein detections occupying the same origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* clades: HLI (including *P. marinus* MED4, solid black circles), LLI (including *P. marinus* NATL2A, solid black circles), LLII/III (including *P. marinus* SS120, solid black circles) and LLIV (including *P. marinus* MIT9313, solid black circles).",
                       "Columns show detections of proteins annotated as Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison, culture growth rate experimental conditions are indicated by horizontal grey lines for depths approximating peak Photosynthetically Active Radiation (PAR; µmol photons m^-2^ s^-1^); and vertical grey lines for imposed [O~2~] (µM).",
                       "Data obtained from the Biological and Chemical Oceanography Data Management Office repository [@saitoPeptidesTheirSpectral2018].",
 .sep = " ")

@@

# This is Fig4.tiff
SS120PARGAM_cap <- glue("**A. Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 (Low-Light (LLII/III) deep ocean clade) vs. photoperiod (h). **",
                    "Rows separate data from levels of imposed dissolved O~2~ concentrations (250 µM, 25 µM and 2.5 µM).",
                    "Columns separate data from 3 levels of peak imposed Photosynthetically Active Radiation (PAR; 30, 90 and 180 µmol photons m^-2^ s^-1^).",
                    "Colours represent the imposed spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",
                    "**B. Contour plot of a Generalized Additive Model (GAM) representing the chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 grown under 660 nm (red) or 450 nm (blue) light.**",
                    "X-axis is photoperiod (h).",
                    "Y-axis is Photosynethetically Active Radiation (PAR, µmol photons m^-2^ s^-1^).",
                   "Top left panel represents the model under 250 µM of O~2~ and red light.",
                    "Bottom left panel represents the model under 25 µM of O~2~ and red light.",
                    "Top right panel represents the model under 250 µM of O~2~ and blue light.",
                    "Bottom right panel represents the model under 25 µM of O~2~ and blue light.",
                    "Legends represent a colour gradient of growth rate from no growth (white) to 1.00 d^-1^ (dark red or dark blue).",
                    "Labeled contour lines indicate the 90%, 50%, and 10% quantiles for achieved growth rate.",

@@

# This is Fig5.tiff
MIT9313PARGAM_cap <- glue("**A. Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MIT9313 (Low-Light (LLIV) deep ocean clade) vs. photoperiod (h). **",
                    "Rows separate data from levels of imposed dissolved O~2~ concentrations (250 µM, 25 µM and 2.5 µM).",
                    "Columns separate data from 3 levels of peak imposed Photosynthetically Active Radiation (PAR; 30, 90 and 180 µmol photons m^-2^ s^-1^).",
                    "Colours represent the imposed spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",
                    "**B. Contour plot of a Generalized Additive Model (GAM) representing the chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MIT9313 grown under 660 nm (red) or 450 nm (blue) light.**",
                    "X-axis is photoperiod (h).",
                    "Y-axis is Photosynthetically Active Radiation (PAR; µmol photons m^-2^ s^-1^).",
                   "Top left panel represents the model under 250 µM of O~2~ and red light.",
                   "Center left panel represents the model under 25 µM of O~2~ and red light.",
                   "Bottom left panel represents the model under 2.5 µM of O~2~ and red light.",
                    "Top right panel represents the model under 250 µM of O~2~ and blue light.",
                    "Center right panel represents the model under 25 µM of O~2~ and blue light.",
                    "Bottom right panel represents the model under 2.5 µM of O~2~ and blue light.",
                    "Legends represent a colour gradient of growth rate from no growth (white) to 1.00 d^-1^ (dark red or dark blue).",
                    "Labeled contour lines indicate the 90%, 50%, and 10% quantiles for achieved growth rate.",

@@

The representative of HLI clade, *P. marinus* MED4, showed no growth
under any 4 h photoperiod treatments, even when a 4 h photoperiod
delivered cumulative diel PUR equivalent to other photoperiod treatments
(Fig \@ref(fig:PhotoperiodPurFitsPlots)A-C). In parallel MED4
showed no growth under 2.5 µM O~2~, no matter the level of diel
cumulative PUR. In contrast, under 250 or 25 µM O~2~, and including
photoperiods greater than 4 h, MED4 growth under blue light was
described by a saturating response of growth rate
[@harrisonPhotosynthesisirradianceRelationshipsPolar1986] to increasing
cumulative diel PUR, with saturation of growth rate achieved around 1.0
x 10^6^ µmol m^-2^d^-1^ (Fig \@ref(fig:BluevsRedPurFitsPlots)A and B),
and no evidence of inhibition of growth at any achieved cumulative diel
PUR. Under the 'artificial' growth treatment of red light, MED4 achieved
more growth per unit diel cumulative PUR (Fig \@ref(fig:BluevsRedPurFitsPlots)A and B), consistent with Murphy *et al*.
[@murphyPhotoinactivationPhotosystemII2017], who showed a lower cost for
growth under red light, for MED4, because red light provokes less
photoinactivation of PSII, than equivalent levels of blue light. For
distinct fits for different photoperiods refer to Fig \@ref(fig:PhotoperiodPurFitsPlots) A-C.

The representative of the LLII/III clade, *P. marinus* SS120 showed
almost no growth under 2.5 µM O~2~ experiments (Fig \@ref(fig:PhotoperiodPurFitsPlots)F). Most 4 h photoperiod treatments of
SS120 also did not grow under 250 µM O~2~ , even when a 4 h photoperiod
delivered cumulative diel PUR equivalent to other photoperiod treatments
(Fig \@ref(fig:PhotoperiodPurFitsPlots)D). Again using
[@harrisonPhotosynthesisirradianceRelationshipsPolar1986], SS120 did not
grow when exposed to more than \~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of
cumulative diel PUR under any spectral waveband or photoperiod
combination, under 250 µM O~2~ (Fig \@ref(fig:PhotoperiodPurFitsPlots)D).

@@

The LLIV clade representative, *P. marinus* MIT9313, under 250 µM O~2~
showed growth rate rising to a plateau by about 5 x 10^5^ µmol photons m^-2^
d^-1^ of cumulative diel PUR. Above about 1.0 x 10^6^ µmol photons m^-2^
d^-1^ of cumulative PUR under 250 µM O~2~, MIT9313 showed full
inhibition of growth, across photoperiods, and spectral wavebands
(Fig \@ref(fig:PhotoperiodPurFitsPlots)G). Under 25 µM O~2~ MIT9313
showed higher growth rates over a wider plateau, with a greatly extended
exploitation of higher cumulative diel PUR, with full growth inhibition
only above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ (Fig \@ref(fig:PhotoperiodPurFitsPlots)H).  Under 2.5 µM O~2~, MIT9313
showed a wider, lower, flatter growth rate response to cumulative diel PUR, with
full growth inhibition only above about 3.5 x 10^6^ µmol photons m^-2^
d^-1^ cumulative diel PUR (Fig \@ref(fig:PhotoperiodPurFitsPlots)I).

@@

*Prochlorococcus marinus* genomes encode 4 FtsH proteins [@mannInvolvementFtsHHomologue2000; @adamFtsHProteasesChloroplasts2005], henceforth referred to
as FtsH1-4, homologs to the characterized FtsH isoforms of the model
freshwater cyanobacterium *Synechocystis sp.* PCC6803, and with presumably
parallel functions (Table \@ref(tab:FtsHTable)). Upon a shift to higher
light, HLI MED4 upregulates expression of FtsH1 and FtsH2
[@bonisteelStrainSpecificDifferences2018], homologs to the
*Synechocystis* *slr*0228 and *slr*1604, implicated in PSII repair [@adamFtsHProteasesChloroplasts2005; @boehmSubunitOrganizationSynechocystis2012]. In
contrast, representative LLIV strain MIT9313 shows no induction of
expression of these FtsH protease isoforms when shifted to high light,
and thus has fewer of these FtsH hexamers serving each photosystem [@bonisteelStrainSpecificDifferences2018].
Transcript analysis demonstrates that MIT9313 expressed primarily FtsH3,
homologous to *Synechocystis* *sll*1463, possibly involved in PSI
biogenesis [@mannInvolvementFtsHHomologue2000; @krynickaFtsH4ProteaseControls2022,
@konikCyanobacterialFtsH4Protease2024]. FtsH3 expression did not
increase in response to light stress in MIT9313
[@bonisteelStrainSpecificDifferences2018]. Through adaptation to steady
low light, clade LLIV *P. marinus* instead allocate resources to
processes other than dynamic regulation of PSII repair.

@@

Ocean detections of proteins mediating protein metabolism support this
interpretation of distinct FtsH function across clades of *P. marinus*.
Ribosome proteins from clade HLI MED4, clade LLI NATL2A, clade LLII/III
SS120 and clade LLIV MIT9313 show generally similar patterns vs. [O~2~]
and depth, a proxy for peak PAR (Fig \@ref(fig:CladeProchloroProteinMetabDepthO2)). FtsH3, inferred to
mediate PSI assembly, likewise shows a similar pattern between MED4 and
MIT9313 (Fig \@ref(fig:CladeProchloroProteinMetabDepthO2)). But only
MED4 shows the presence of the FtsH1 & FtsH2 isoforms inferred to
mediate PSII repair, and then only in near-surface samples subject to
higher light levels. Furthermore, even though MIT9313 grows (Fig \@ref(fig:MIT9313PARGAM)B), and is detected in the ocean at low [O~2~]
(Fig \@ref(fig:CladeProchloroPhotosynthDepthO2)), no FtsH from
MIT9313 is detected at low [O~2~] (Fig \@ref(fig:CladeProchloroProteinMetabDepthO2)), suggesting limited
requirement for protein turnover under low [O~2~].

@@

# This is Fig7.tiff
CladeProchloroProteinMetabDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* protein metabolism complexes.**", 
                       "Protein detections (circles) are plotted vs. O~2~ (µM) (x-axis) and depth (m) (y-axis) at sample origin with a 15% offset to seperate protein detections occupying the same origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* clades: HLI (including *P. marinus* MED4, solid black circles), LLI (including *P. marinus* NATL2A, solid black circles), LLII/III (including *P. marinus* SS120, solid black circles) and LLIV (including *P. marinus* MIT9313, solid black circles).",
                       "Columns show detections of proteins annotated as FtsH Protease Complexes (FtsH1, FtsH2, FtsH3) or the Ribosome.",
                       "For comparison, culture growth rate experimental conditions are indicated by horizontal grey lines for depths approximating peak Photosynthetically Active Radiation (PAR; µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from the Biological and Chemical Oceanography Data Management Office repository [@saitoPeptidesTheirSpectral2018].",
 .sep = " ")

@@

XXXXX To gain insights into the growth rate responses across strains, particularly the responses to different oxygen levels [O~2~], we used a subset of the dataset of annotated phytoplankton genomes from Omar *et al*. [@omarAnnotationGenesEncoding2023]. We extracted the genes encoding enzymes involved in DNA repair found in at least one *P.marinus* strain in this study *P.marinus* strain (Fig \@ref(fig:DNARepairFig). XXXX

>>>>

*Prochlorococcus marinus*, the smallest picocyanobacterium, comprises
multiple clades with distinct niches, currently across tropical and sub-tropical
oligotrophic ocean regions, including Oxygen Minimum Zones. Ocean
warming may open permissive temperatures in new, poleward photic
regimes, along with expanded Oxygen Minimum Zones. We used ocean metaproteomic
data to guide testing of *Prochlorococcus marinus* growth across a matrix of peak irradiances,
photoperiods, spectral bands and dissolved oxygen. MED4 from Clade HLI
requires greater than 4 h photoperiod, grows at 25 µmol O~2~ L^-1^ and
above, and exploits high cumulative diel photon doses. MED4, however, relies upon an alternative oxidase to
balance electron transport, which may exclude it from growth under our
lowest, 2.5 µmol O~2~ L^-1^, condition. SS120 from clade LLII/III is
restricted to low light under full 250 µmol O~2~ L^-1^, shows expanded
light exploitation under 25 µmol O~2~ L^-1^, but is excluded from growth
under 2.5 µmol O~2~ L^-1^. Intermediate oxygen suppresses the cost of
PSII photoinactivation, and possibly the enzymatic production of H~2~O~2~ in SS120,
which has limitations on genomic capacity for PSII and DNA repair. MIT9313 from
Clade LLIV is restricted to low blue irradiance under 250 µmol O~2~
L^-1^, but exploits much higher irradiance under red light, or under
lower O~2~ concentrations, conditions which slow photoinactivation of
PSII and production of reactive oxygen species. In warming oceans, range expansions and competition among clades will be governed not simply by light levels. Short photoperiods governed by latitude, temperate winter, and depth attenuation of light, will exclude clade HLI (including MED4). In contrast, clade LLII/III (including SS120), and especially clade LLIV (including MIT9313), may actually exploit higher light niches under expanding OMZ conditions, where low O~2~ relieves the stresses of oxidation stress and PSII photoinhibition.

@@

Low-Light clades thrive in deeper ocean waters, extending beyond 200 m
in depth [@partenskyProchlorococcusMarinePhotosynthetic1999], where only
\~1% of the surface irradiance penetrates, primarily in the blue (450
nm) to green (520 nm) spectral range
[@holtropVibrationalModesWater2021]. Clade LLI includes cultured strain
NATL2A, which prefers moderate irradiances typical of between 30 and 100
m depth. Clades LLII and LLIII, including cultured strain SS120, are
grouped together as early branching phylogenetic lineage in
the *P. marinus* radiation, with a preference for low light. Clade
LLIV, including cultured strain MIT9313, falls near the base of the
*P. marinus* radiation, and has been characterized by preference for low
light, typical of depths from 120 m to 200 m
[@partenskyProchlorococcusMarinePhotosynthetic1999]. Clade LLIV members are,
as yet, the only cultured *P. marinus* strains to have been found in Oxygen Minimum
Zones (OMZ). Some, as yet, uncultured *P. marinus* strains in clades LLV
and LLVI also thrive in OMZ of the subtropical Atlantic and Pacific
Oceans, where dissolved oxygen concentrations [O~2~] can be less than 20
µM [@ulloaCyanobacteriumProchlorococcusHas2021;
@lavinNovelLineagesProchlorococcus2010;
@goerickeNovelNicheProchlorococcus2000;
@garcia-robledoCrypticOxygenCycling2017;
@johnsonEnergeticsGrowthKinetics1999]. *Prochlorococcus marinus* LL
ecotypes may indeed dominate the phytoplankton within OMZ
[@lavinNovelLineagesProchlorococcus2010;
@ulloaCyanobacteriumProchlorococcusHas2021;
@partenskyProchlorococcusAdvantagesLimits2010], where they may be net
O~2~ consumers [@partenskyComparisonPhotosyntheticPerformances2018].

@@

## Detection of proteins from *Prochlorococcus marinus* Clades across O~2~ and light niches in the ocean {.unnumbered}

Proteins from *P. marinus* were detected across depths and oxygen concentrations in the OPP data set [@saitoPeptidesTheirSpectral2018]. We focused our analyses on core photosynthetic protein complexes, annotated as coming from clades HLI (2 strains, including strain MED4 ); LLI (2 strains, including strain NATL2A); LLII/III (2 strains, including strain SS120) and LLIV (2 strains, including MIT9313) (Fig \@ref(fig:CladeProchloroPhotosynthDepthO2)), as a function of depth (a
proxy for light level) and measured [O~2~] at the sampling locations. Photosynthetic complexes
from clades HLI and LLI were detected throughout the water column, although predominately at stations with high [O~2~]. Though present in the genome of MED4, the absence of proteins annotated as RUBISCO complex for clade HLI, compared to annotated
detections of RUBISCO across the other three clades, suggests
limitations in the annotation process assigning highly conserved protein
sequences to clades. Notwithstanding limitations on assignments of proteins to clades, detections of the abundant carbon fixation complex RUBISCO derived from clades HLI, LLI and LLII/III were notably absent from stations at low [O~2~], suggesting limited capacity for carbon fixation by clades HLI,  LLI and LLII/LLIII under low [O~2~] habitats. Clade LLII/LLIII
photosynthetic complexes were detected throughout the sampled ranges of depth/light and [O~2~], with  detections of proteins at both high and low [O~2~], compared to other strains. Complexes from LLIV (including strain MIT9313) were detected across the depth/light and [O~2~] ranges, with more detections at deeper, darker depths and at low [O~2~], compared to other clades.  We note that detections of photosynthetic proteins from strain MIT9313 (dark symbols) show a particular bias towards low [O~2~] sampling stations, compared to the wider detections of clade LLIV. Our analyses utilized a proteomics dataset; however, alternate approaches, such as metagenomics or metatranscriptomics, could have been employed to analyze ecotype abundances using the TARA Oceans [@pesantOpenScienceResources2015] or other datasets.

@@

# This is Fig2.tiff
CladeProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections (circles) are plotted vs. O~2~ (µM) (x-axis) and depth (m) (y-axis) at sample origins, with 'jitter' offsets up to 15% of full axes scales, to visualize over-laid data points.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* clades: HLI (including *P. marinus* MED4, solid black circles), LLI (including *P. marinus* NATL2A, solid black circles), LLII/III (including *P. marinus* SS120, solid black circles) and LLIV (including *P. marinus* MIT9313, solid black circles).",
                       "Columns show detections of proteins annotated as Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison, experimental conditions for culture growth rate determinations are indicated by horizontal grey lines for depths, approximating peak Photosynthetically Active Radiation (PAR; µmol photons m^-2^ s^-1^); and vertical grey lines for imposed [O~2~] (µM).",
                       "Data obtained from the Biological and Chemical Oceanography Data Management Office repository [@saitoPeptidesTheirSpectral2018].",
 .sep = " ")

@@

# This is Fig4.tiff
SS120PARGAM_cap <- glue("**A. Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 (Low-Light, deep ocean clade LLII/III) vs. photoperiod (h). **",
                    "Rows separate data from levels of imposed dissolved O~2~ concentrations (250 µM, 25 µM and 2.5 µM).",
                    "Columns separate data from 3 levels of peak imposed Photosynthetically Active Radiation (PAR; 30, 90 and 180 µmol photons m^-2^ s^-1^).",
                    "Colours represent the imposed spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",
                    "**B. Contour plot of a Generalized Additive Model (GAM) representing the chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 grown under 660 nm (red) or 450 nm (blue) light.**",
                    "X-axis is photoperiod (h).",
                    "Y-axis is Photosynethetically Active Radiation (PAR, µmol photons m^-2^ s^-1^).",
                   "Top left panel represents the model under 250 µM of O~2~ and red light.",
                    "Bottom left panel represents the model under 25 µM of O~2~ and red light.",
                    "Top right panel represents the model under 250 µM of O~2~ and blue light.",
                    "Bottom right panel represents the model under 25 µM of O~2~ and blue light.",
                    "Legends represent a colour gradient of growth rate from no growth (white) to 1.00 d^-1^ (dark red or dark blue).",
                    "Labeled contour lines indicate the 90%, 50%, and 10% quantiles for achieved growth rate.",

@@

# This is Fig5.tiff
MIT9313PARGAM_cap <- glue("**A. Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MIT9313 (Low-Light, deep ocean clade LLIV) vs. photoperiod (h). **",
                    "Rows separate data from levels of imposed dissolved O~2~ concentrations (250 µM, 25 µM and 2.5 µM).",
                    "Columns separate data from 3 levels of peak imposed Photosynthetically Active Radiation (PAR; 30, 90 and 180 µmol photons m^-2^ s^-1^).",
                    "Colours represent the imposed spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",
                    "**B. Contour plot of a Generalized Additive Model (GAM) representing the chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MIT9313 grown under 660 nm (red) or 450 nm (blue) light.**",
                    "X-axis is photoperiod (h).",
                    "Y-axis is Photosynthetically Active Radiation (PAR; µmol photons m^-2^ s^-1^).",
                   "Top left panel represents the model under 250 µM of O~2~ and red light.",
                   "Center left panel represents the model under 25 µM of O~2~ and red light.",
                   "Bottom left panel represents the model under 2.5 µM of O~2~ and red light.",
                    "Top right panel represents the model under 250 µM of O~2~ and blue light.",
                    "Center right panel represents the model under 25 µM of O~2~ and blue light.",
                    "Bottom right panel represents the model under 2.5 µM of O~2~ and blue light.",
                    "Legends represent a colour gradient of growth rate from no growth (white) to 1.00 d^-1^ (dark red or dark blue).",
                    "Labeled contour lines indicate the 90%, 50%, and 10% quantiles for achieved growth rate.",

@@

The representative of clade HLI, *P. marinus* MED4, showed no growth
under any 4 h photoperiod treatments, even when a 4 h photoperiod
delivered cumulative diel PUR equivalent to other photoperiod treatments
(Fig \@ref(fig:PhotoperiodPurFitsPlots)A-C). In parallel MED4
showed no growth under 2.5 µM O~2~, no matter the level of diel
cumulative PUR. In contrast, under 250 or 25 µM O~2~, and including
photoperiods greater than 4 h, MED4 growth under blue light was
described by a saturating response of growth rate
[@harrisonPhotosynthesisirradianceRelationshipsPolar1986] to increasing
cumulative diel PUR, with saturation of growth rate achieved around 1.0
x 10^6^ µmol m^-2^d^-1^ (Fig \@ref(fig:BluevsRedPurFitsPlots)A and B),
and no evidence of inhibition of growth at any achieved cumulative diel
PUR. Under the 'artificial' growth treatment of red light, MED4 achieved
more growth per unit diel cumulative PUR (Fig \@ref(fig:BluevsRedPurFitsPlots)A and B), consistent with Murphy *et al*.
[@murphyPhotoinactivationPhotosystemII2017], who showed a lower cost for
growth under red light, for MED4, because red light provokes less
photoinactivation of PSII, than equivalent levels of blue light. For
distinct fits for different photoperiods refer to Fig \@ref(fig:PhotoperiodPurFitsPlots) A-C.

The representative of clade LLII/III, *P. marinus* SS120 showed
almost no growth under 2.5 µM O~2~ experiments (Fig \@ref(fig:PhotoperiodPurFitsPlots)F). Most 4 h photoperiod treatments of
SS120 also did not grow under 250 µM O~2~ , even when a 4 h photoperiod
delivered cumulative diel PUR equivalent to other photoperiod treatments
(Fig \@ref(fig:PhotoperiodPurFitsPlots)D). Again using
[@harrisonPhotosynthesisirradianceRelationshipsPolar1986], SS120 did not
grow when exposed to more than \~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of
cumulative diel PUR under any spectral waveband or photoperiod
combination, under 250 µM O~2~ (Fig \@ref(fig:PhotoperiodPurFitsPlots)D).

@@

The clade LLIV representative, *P. marinus* MIT9313, under 250 µM O~2~
showed growth rate rising to a plateau by about 5 x 10^5^ µmol photons m^-2^
d^-1^ of cumulative diel PUR. Above about 1.0 x 10^6^ µmol photons m^-2^
d^-1^ of cumulative PUR under 250 µM O~2~, MIT9313 showed full
inhibition of growth, across photoperiods, and spectral wavebands
(Fig \@ref(fig:PhotoperiodPurFitsPlots)G). Under 25 µM O~2~ MIT9313
showed higher growth rates over a wider plateau, with a greatly extended
exploitation of higher cumulative diel PUR, with full growth inhibition
only above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ (Fig \@ref(fig:PhotoperiodPurFitsPlots)H).  Under 2.5 µM O~2~, MIT9313
showed a wider, lower, flatter growth rate response to cumulative diel PUR, with
full growth inhibition only above about 3.5 x 10^6^ µmol photons m^-2^
d^-1^ cumulative diel PUR (Fig \@ref(fig:PhotoperiodPurFitsPlots)I).

@@

*Prochlorococcus marinus* genomes encode 4 FtsH proteins [@mannInvolvementFtsHHomologue2000; @adamFtsHProteasesChloroplasts2005], henceforth referred to as FtsH1-4, homologs to the characterized FtsH isoforms of the model freshwater cyanobacterium *Synechocystis sp.* PCC6803, and with presumably parallel functions (Table \@ref(tab:FtsHTable)). Upon a shift to higher light, clade HLI MED4 upregulates expression of FtsH1 and FtsH2
[@bonisteelStrainSpecificDifferences2018], homologs to the
*Synechocystis* *slr*0228 and *slr*1604, implicated in PSII repair [@adamFtsHProteasesChloroplasts2005; @boehmSubunitOrganizationSynechocystis2012]. In
contrast, representative clade LLIV strain MIT9313 shows no induction of
expression of these FtsH protease isoforms when shifted to high light,
and thus has fewer of these FtsH hexamers serving each photosystem [@bonisteelStrainSpecificDifferences2018].
Transcript analysis demonstrates that MIT9313 expressed primarily FtsH3,
homologous to *Synechocystis* *sll*1463, possibly involved in PSI
biogenesis [@mannInvolvementFtsHHomologue2000; @krynickaFtsH4ProteaseControls2022,
@konikCyanobacterialFtsH4Protease2024]. FtsH3 expression did not
increase in response to light stress in MIT9313
[@bonisteelStrainSpecificDifferences2018]. Through adaptation to steady
low light, clade LLIV *P. marinus* instead allocates resources to
processes other than dynamic regulation of PSII repair.

@@

Ocean detections of proteins [@saitoPeptidesTheirSpectral2018] mediating protein metabolism support this interpretation of distinct FtsH function across clades of *P. marinus*  (Fig \@ref(fig:CladeProchloroProteinMetabDepthO2)). Ribosome proteins annotated as from clade HLI (including MED4) were not detected at stations with low [O~2~], and near surface detections of ribosomes from clade HLI were far more frequent than detections at depth. Ribosome proteins from clade LLI (including NATL2A); clade LLII/III
(including SS120) and clade LLIV (including MIT9313) show generally similar patterns vs. [O~2~] and depth, with at least some detections from stations with low [O~2~]. FtsH3, inferred to
mediate PSI assembly, likewise shows a similar pattern between clade HLI and
clade LLIV (Fig \@ref(fig:CladeProchloroProteinMetabDepthO2)). Only
clade HLI shows the presence of FtsH1 isoforms, inferred to partner with FtsH2, to
mediate PSII repair, and then only in near-surface samples subject to
higher light levels. Furthermore, even though clade IV grows (Fig \@ref(fig:MIT9313PARGAM)B), and is detected in the ocean at low [O~2~]
(Fig \@ref(fig:CladeProchloroPhotosynthDepthO2)), no FtsH from
clade IV, nor indeed FtsH from any clade, is detected at low [O~2~] (Fig \@ref(fig:CladeProchloroProteinMetabDepthO2)), consistent with limited
requirements for protein turnover under low [O~2~].

@@

# This is Fig7.tiff
CladeProchloroProteinMetabDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* protein metabolism complexes.**", 
                       "Protein detections (circles) are plotted vs. O~2~ (µM) (x-axis) and depth (m) (y-axis) at sample origin with a 15% offset to seperate protein detections occupying the same origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* clades: HLI (including *P. marinus* MED4, solid black circles), LLI (including *P. marinus* NATL2A, solid black circles), LLII/III (including *P. marinus* SS120, solid black circles) and LLIV (including *P. marinus* MIT9313, solid black circles).",
                       "Columns show detections of proteins annotated as FtsH Protease Complexes (FtsH1, FtsH2, FtsH3) or the Ribosome.",
                       "For comparison, experimental conditions for culture growth rate determinations are indicated by horizontal grey lines for depths approximating peak Photosynthetically Active Radiation (PAR; µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from the Biological and Chemical Oceanography Data Management Office repository [@saitoPeptidesTheirSpectral2018].",
 .sep = " ")

@@

To gain insights into the distinct growth rate responses across strains, particularly the responses to different oxygen levels [O~2~], we used a dataset of annotated phytoplankton genomes from Omar *et al*. [@omarAnnotationGenesEncoding2023]. We extracted those genes encoding enzymes involved in DNA repair present in at least one of the *P.marinus* study strains (Fig \@ref(fig:DNARepairFig).
